---
title: 2_filter_data
author: Johannes B. Gruber
date: today
format:
  html:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    embed-resources: true
---

# Introduction

Our goal is to find out whether we can identify the weaponization of LGBTQ+ rights in party communication.
For this, we download data from the Guardian API.
We want articles that mention LGBTQ+ rights and a party or a politician affiliated with a party.

## Packages

I use httr2 to pull articles from the Guardian API.

```{r setup}
#| include: false
library(tidyverse)
library(tidytext)
library(glue)
library(httr2)
start <- Sys.time() # note start time for later
prototype <- FALSE
```

## Data

```{r}
articles_guardian <- readRDS("data/articles_guardian.rds") |> 
  filter(web_publication_date > "1999-01-01",
         !duplicated(id))

lgbtq_terms <- readLines("lgbtq-dict.txt") |> 
  str_remove("#.*") |> 
  str_trim() |> 
  stringi::stri_remove_empty()

parties <- readLines("party-dict.txt")

parties_uk <- parties[1:str_which(parties, fixed("# US (since 1990)"))] |> 
  str_remove("#.*") |> 
  str_trim() |> 
  stringi::stri_remove_empty()
parties_us <- parties[str_which(parties, fixed("# US (since 1990)")):length(parties)] |> 
  str_remove("#.*") |> 
  str_trim() |> 
  stringi::stri_remove_empty()

dict <- tibble(
  word = c(lgbtq_terms, parties_uk, parties_us),
  set = c(rep("lgbtq", length(lgbtq_terms)),
          rep("parties_uk", length(parties_uk)),
          rep("parties_us", length(parties_us)))
) |> 
  distinct(word, .keep_all = TRUE)
dict

if (prototype) {
  set.seed(1)
  articles_guardian <- articles_guardian |> 
    sample_n(500)
}
```


# Main

I retain only paragraphs that feature both a keyword from the LGBTQ list and one from the party list.

```{r}
paragraphs <- articles_guardian |> 
  unnest_paragraphs(output = text, input = text, to_lower = FALSE) |>
  group_by(id) |> 
  mutate(par_id = row_number(), .after = id) |> 
  ungroup() |> 
  unnest_tokens(output = word, input = text, to_lower = FALSE, drop = FALSE) |> 
  inner_join(dict, by = "word") |> 
  group_by(id, par_id) |> 
  summarise(
    text = head(text, 1),
    word = list(word),
    set = list(sort(unique(set))), 
    .groups = "drop"
  ) |> 
  filter(map_lgl(set, \(x) "lgbtq" %in% x && any(x %in% c("parties_uk", "parties_us"))))
```

This retains:

```{r}
par_count <- paragraphs |> 
  count(id)

tribble(
  ~"", ~Number,
  "total", scales::comma(nrow(paragraphs)),
  "avg. per article", scales::comma(mean(par_count$n), accuracy = 0.01),
  "article max.",  scales::comma(max(par_count$n)),
  "article min.",  scales::comma(min(par_count$n)),
  "no hit", scales::comma(sum(!articles_guardian$id %in% paragraphs$id)),
  "no hit pct", scales::percent(sum(!articles_guardian$id %in% paragraphs$id) / nrow(articles_guardian))
)
```

```{r}
art_counts <- articles_guardian |> 
  filter(web_publication_date > "1999-01-01") |> 
  count(date = lubridate::floor_date(web_publication_date, "month")) |> 
  mutate(what = "articles")

par_counts <- paragraphs |> 
  left_join(articles_guardian, by = "id") |> 
  count(date = lubridate::floor_date(web_publication_date, "month")) |> 
  mutate(what = "selected paragraphs")

freq_plot <- bind_rows(art_counts, par_counts) |> 
  ggplot(aes(x = date, y = n, colour = what)) +
  geom_line() +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme_minimal() +
  scale_color_manual(values = c("articles" = "#2D2926", 
                                "selected paragraphs" = "#970014")) +
  scale_x_datetime(date_breaks = "3 years", date_labels = "%Y") +
  theme(legend.position = "inside", legend.position.inside = c(0.1, 0.9))
ggsave("figures/fig1.png", freq_plot, width = 16, height = 11, units = "cm")
freq_plot
```

```{r}
freq_plot_alt <- par_counts |> 
  ggplot(aes(x = date, y = n, colour = what)) +
  geom_line() +
  labs(x = NULL, y = NULL, colour = NULL) +
  theme_minimal() +
  scale_color_manual(values = c("articles" = "#2D2926", 
                                "selected paragraphs" = "#970014")) +
  scale_x_datetime(date_breaks = "3 years", date_labels = "%Y") +
  theme(legend.position = "inside", legend.position.inside = c(0.1, 0.9))
ggsave("figures/fig1_alt.png", freq_plot_alt, width = 16, height = 11, units = "cm")
freq_plot
```


# Wrap-up

```{r}
saveRDS(paragraphs, "data/lgbtq_party_paragraphs.rds")
```


Afterwards we get some information which is important to reproduce the report.

```{r}
sessionInfo()
Sys.time()
# note how long the script takes to (re-)run
Sys.time() - start
```
