---
title: 4_analysis
author: Johannes B. Gruber
date: today
format:
  html:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    embed-resources: true
---

# Introduction


## Packages

```{r setup}
#| include: false
library(tidyverse)
library(jsonlite)
library(patchwork)
start <- Sys.time() # note start time for later
prototype <- TRUE
```

## Data

```{r}
articles_guardian <- readRDS("data/articles_guardian.rds") |> 
  filter(web_publication_date > "1999-01-01",
         !duplicated(id))
paragraphs_annotated <- readRDS("data/lgbtq_party_paragraphs_annotated.rds")
# I cleaned up the party names by hand to make them easier to plot
paragraphs_annotated_party <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/13qfw2IR_ovqN7SH9FR9eP-ZAJZpEYMhSAw5396Wzkr8/edit?usp=sharing") |> 
  select(id, par_id, `5_party`, `5_party_clean`)
```


# Main

Some data wrangling and joining to get a nice clean analysis dataset.

```{r}
paragraphs_annotated_tidy <- paragraphs_annotated |> 
  rowwise() |> 
  mutate(annotation = str_extract(annotation, "\\{.+?\\}"),
         annotation = list(fromJSON(annotation))) |> 
  unnest_wider(annotation) |> 
  left_join(articles_guardian |> select(id, web_publication_date),
            by = "id") |> 
  left_join(paragraphs_annotated_party, by = c("id", "par_id", "5_party")) |> 
  # clean up variable
  mutate(evaluation_lgbtq_clean = `2_evaluation_lgbtq`,
         evaluation_lgbtq_clean = str_remove(evaluation_lgbtq_clean, " \\(.*"),
         evaluation_lgbtq_clean = ifelse(evaluation_lgbtq_clean != "NA", evaluation_lgbtq_clean, "not mentioned"),
         evaluation_lgbtq_clean = ifelse(`1_lgbtq` == "Yes", evaluation_lgbtq_clean, "not mentioned"),
         evaluation_lgbtq_clean = ifelse(evaluation_lgbtq_clean != "concerned/neutral", evaluation_lgbtq_clean, "positive"),
         evaluation_lgbtq_clean = fct_relevel(evaluation_lgbtq_clean, "not mentioned", "positive", "neutral", "negative")) |> 
  mutate(evaluation_minority_clean = `4_evaluation_minority`,
         evaluation_minority_clean = str_remove(evaluation_minority_clean, " \\(.*"),
         evaluation_minority_clean = ifelse(evaluation_minority_clean != "NA", evaluation_minority_clean, "not mentioned"),
         evaluation_minority_clean = ifelse(`1_lgbtq` == "Yes", evaluation_minority_clean, "not mentioned"),
         evaluation_minority_clean = fct_relevel(evaluation_minority_clean, "not mentioned", "positive", "neutral", "negative"))
```

## Party plots

And some more wrangling for the party plots. Since there are a lot of parties, I collapse everything but the most frquent ones into "others". I also collapse Reform UK and UKIP into one.

```{r}
top_parties <- paragraphs_annotated_tidy |> 
  separate_longer_delim(`5_party_clean`, ", ") |> 
  count(`5_party_clean`, sort = TRUE)

paragraphs_annotated_tidy_clean <- paragraphs_annotated_tidy |> 
  separate_longer_delim(`5_party_clean`, ", ") |> 
  mutate(party_clean = recode_factor(
    `5_party_clean`,
    "Labour"            = "Labour"           ,
    "Republicans"       = "Republicans"      ,
    "Democrats"         = "Democrats"        ,
    "Conservatives"     = "Conservatives"    ,
    "Green Party"       = "Green Party"      ,
    "DUP"               = "DUP"              ,
    "Liberal Democrats" = "Liberal Democrats",
    "Labor"             = "Labor"            ,
    "SNP"               = "SNP"              ,
    "Liberals"          = "Liberals"         ,
    "UKIP/Reform"       = "UKIP"             ,
    "Liberal Party"     = "Liberal Party"    ,
    "Fine Gael"         = "Fine Gael"        ,
    "UKIP/Reform"       = "Reform UK"        ,
    .default = "other",
    .missing = "No mention"
  ), 
  party_clean = fct_rev(fct_infreq(party_clean)))
```

First plot to show how often parties evaluate LGBTQ+ issues or rights as positive/neutral/negative:

```{r}
fig2_lgbtq_sentiment <- paragraphs_annotated_tidy_clean |> 
  count(party_clean, evaluation_lgbtq_clean, sort = TRUE) |> 
  ggplot(aes(x = n, y = party_clean, fill = evaluation_lgbtq_clean)) +
  geom_col() +
  scale_fill_manual(values = c(
    "negative" = "#fc4e51",
    "neutral" = "#ffb347",
    "positive" = "#77c97d",
    "not mentioned" = "#d3d3d3"
  )) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal() +
  theme(legend.position = "inside", legend.position.inside = c(0.8, 0.2))
ggsave("figures/fig2.png", fig2_lgbtq_sentiment, width = 16, height = 11, units = "cm")
fig2_lgbtq_sentiment
```

And the same with immigrants and ethnic minority issues/rights:

```{r}
fig3_minority_sentiment <- paragraphs_annotated_tidy_clean |> 
  count(party_clean, evaluation_minority_clean, sort = TRUE) |> 
  ggplot(aes(x = n, y = party_clean, fill = evaluation_minority_clean)) +
  geom_col() +
  scale_fill_manual(values = c(
    "negative" = "#fc4e51",
    "neutral" = "#ffb347",
    "positive" = "#77c97d",
    "not mentioned" = "#d3d3d3"
  )) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal() +
  theme(legend.position = "inside", legend.position.inside = c(0.8, 0.2))
ggsave("figures/fig3.png", fig3_minority_sentiment, width = 16, height = 11, units = "cm")
fig3_minority_sentiment
```

Since they are so similar, we can collapse them into one plot:

```{r}
fig2_alt <- fig2_lgbtq_sentiment + fig3_minority_sentiment + plot_layout(axes = "collect")
ggsave("figures/fig2_alt.png", fig2_alt, width = 16, height = 11, units = "cm")
fig2_alt
```

And to look a bit closer I want to know how often parties see LGBTQ+ as positive and minorities as negative and so on:

```{r}
colfunc <- colorRampPalette(c("#5a5a5a", "#C6C6C6"))
greys <- colfunc(10)
colfunc <- colorRampPalette(c("#77c97d", "#fc4e51"))
cols <- colfunc(4)

fig4_evaluation_combined <- paragraphs_annotated_tidy_clean |>
  count(
    party_clean, evaluation_lgbtq_clean,
    evaluation_minority_clean
  ) |>
  unite(
    col = "Evaluation",
    evaluation_lgbtq_clean, evaluation_minority_clean, sep = "/"
  ) |>
  mutate(Evaluation = factor(Evaluation, levels = rev(c(
    # highlight
    "positive/positive",
    "positive/negative",
    "negative/positive",
    "negative/negative",
    # not highlight
    "positive/neutral",
    "neutral/positive",
    "neutral/neutral",
    "negative/neutral",
    "neutral/negative",
    "positive/not mentioned",
    "neutral/not mentioned",
    "negative/not mentioned",
    "not mentioned/negative",
    "not mentioned/not mentioned"
  )))) |>
  ggplot(aes(x = n, y = party_clean, fill = Evaluation)) +
  geom_col() +
  scale_fill_manual(values = c(
    # highlight
    "positive/positive" = "#77c97d",
    "positive/negative" = "#ffbf47",
    "negative/positive" = "#ffe347",
    "negative/negative" = "#fc4e51",
    # not highlight
    "positive/neutral" = greys[1],
    "neutral/positive" = greys[2],
    "neutral/neutral" = greys[3],
    "negative/neutral" = greys[4],
    "neutral/negative" = greys[5],
    "positive/not mentioned" = greys[6],
    "neutral/not mentioned" = greys[7],
    "negative/not mentioned" = greys[8],
    "not mentioned/negative" = greys[9],
    "not mentioned/not mentioned" = greys[10]
  )) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal()
ggsave("figures/fig4.png", fig4_evaluation_combined, width = 16, height = 11, units = "cm")
fig4_evaluation_combined
```

```{r}
fig4_alt_evaluation_combined <- paragraphs_annotated_tidy_clean |>
  group_by(
    party_clean, evaluation_lgbtq_clean, evaluation_minority_clean
  ) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(party_clean) |> 
  mutate(freq = n / sum(n)) |>
  unite(
    col = "Evaluation",
    evaluation_lgbtq_clean, evaluation_minority_clean, sep = "/"
  ) |>
  mutate(Evaluation = factor(Evaluation, levels = rev(c(
    # highlight
    "positive/positive",
    "positive/negative",
    "negative/positive",
    "negative/negative",
    # not highlight
    "positive/neutral",
    "neutral/positive",
    "neutral/neutral",
    "negative/neutral",
    "neutral/negative",
    "positive/not mentioned",
    "neutral/not mentioned",
    "negative/not mentioned",
    "not mentioned/negative",
    "not mentioned/not mentioned"
  )))) |>
  ggplot(aes(x = freq, y = party_clean, fill = Evaluation)) +
  geom_col() +
  scale_fill_manual(values = c(
    # highlight
    "positive/positive" = "#77c97d",
    "positive/negative" = "#ffbf47",
    "negative/positive" = "#ffe347",
    "negative/negative" = "#fc4e51",
    # not highlight
    "positive/neutral" = greys[1],
    "neutral/positive" = greys[2],
    "neutral/neutral" = greys[3],
    "negative/neutral" = greys[4],
    "neutral/negative" = greys[5],
    "positive/not mentioned" = greys[6],
    "neutral/not mentioned" = greys[7],
    "negative/not mentioned" = greys[8],
    "not mentioned/negative" = greys[9],
    "not mentioned/not mentioned" = greys[10]
  )) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal()
ggsave("figures/fig4_alt.png", fig4_alt_evaluation_combined, width = 16, height = 11, units = "cm")
fig4_alt_evaluation_combined
```

## Time plots

Statment frequency over time

```{r}
fig5_evaluation_over_time <- paragraphs_annotated_tidy_clean |> 
  count(date = lubridate::floor_date(web_publication_date, "year"), evaluation_lgbtq_clean, evaluation_minority_clean) |>
  unite(col = "Evaluation", evaluation_lgbtq_clean, evaluation_minority_clean, sep = "/") |> 
  mutate(Evaluation = factor(Evaluation, levels = rev(c(
    # highlight
    "positive/positive",
    "positive/negative",
    "negative/positive",
    "negative/negative",
    # not highlight
    "positive/neutral",
    "neutral/positive",
    "neutral/neutral",
    "negative/neutral",
    "neutral/negative",
    "positive/not mentioned",
    "neutral/not mentioned",
    "negative/not mentioned",
    "not mentioned/negative",
    "not mentioned/not mentioned"
  )))) |> 
  ggplot(aes(x = date, y = n, colour = Evaluation)) +
  geom_line() + 
  scale_colour_manual(values = c(
    # highlight
    "positive/positive" = "#77c97d",
    "positive/negative" = "#ffbf47",
    "negative/positive" = "#ffe347",
    "negative/negative" = "#fc4e51",
    # not highlight
    "positive/neutral" = greys[1],
    "neutral/positive" = greys[2],
    "neutral/neutral" = greys[3],
    "negative/neutral" = greys[4],
    "neutral/negative" = greys[5],
    "positive/not mentioned" = greys[6],
    "neutral/not mentioned" = greys[7],
    "negative/not mentioned" = greys[8],
    "not mentioned/negative" = greys[9],
    "not mentioned/not mentioned" = greys[10]
  )) +
  labs(x = NULL, y = NULL, fill = NULL, colour = NULL) +
  theme_minimal()
ggsave("figures/fig5.png", fig5_evaluation_over_time, width = 16, height = 11, units = "cm")
fig5_evaluation_over_time
```

Alternative 1: remove neutral and not mentioend:

```{r}
fig5_evaluation_over_time_alt <- paragraphs_annotated_tidy_clean |>
  count(date = lubridate::floor_date(web_publication_date, "year"), evaluation_lgbtq_clean, evaluation_minority_clean) |>
  filter(
    evaluation_lgbtq_clean != "not mentioned",
    evaluation_lgbtq_clean != "neutral",
    evaluation_minority_clean != "not mentioned",
    evaluation_minority_clean != "neutral",
  ) |>
  unite(col = "Evaluation", evaluation_lgbtq_clean, evaluation_minority_clean, sep = "/") |>
  ggplot(aes(x = date, y = n, colour = Evaluation)) +
  geom_line() +
  scale_colour_manual(values = c(
    # highlight
    "positive/positive" = "#77c97d",
    "positive/negative" = "#ffbf47",
    "negative/positive" = "#ffe347",
    "negative/negative" = "#fc4e51"
  )) +
  labs(x = NULL, y = NULL, fill = NULL, colour = NULL) +
  theme_minimal()
ggsave("figures/fig5_alt.png", fig5_evaluation_over_time_alt, width = 16, height = 11, units = "cm")
fig5_evaluation_over_time_alt
```

Alternative 2: make barplot instead

```{r}
fig5_evaluation_over_time_alt2 <- paragraphs_annotated_tidy_clean |>
  count(date = lubridate::floor_date(web_publication_date, "year"), evaluation_lgbtq_clean, evaluation_minority_clean) |>
  filter(
    evaluation_lgbtq_clean != "not mentioned",
    evaluation_lgbtq_clean != "neutral",
    evaluation_minority_clean != "not mentioned",
    evaluation_minority_clean != "neutral",
  ) |>
  unite(col = "Evaluation", evaluation_lgbtq_clean, evaluation_minority_clean, sep = "/") |>
  ggplot(aes(x = date, y = n, fill = Evaluation)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c(
    # highlight
    "positive/positive" = "#77c97d",
    "positive/negative" = "#ffbf47",
    "negative/positive" = "#ffe347",
    "negative/negative" = "#fc4e51"
  )) +
  labs(x = NULL, y = NULL, fill = NULL, colour = NULL) +
  theme_minimal()
ggsave("figures/fig5_alt2.png", fig5_evaluation_over_time_alt2, width = 16, height = 11, units = "cm")
fig5_evaluation_over_time_alt2
```

Alternative 3: instead of yearly frequencies, show smoothed trend:

```{r}
fig5_evaluation_over_time_alt3 <- paragraphs_annotated_tidy_clean |>
  filter(web_publication_date < "2024-01-01") |> 
  count(date = lubridate::floor_date(web_publication_date, "year"), evaluation_lgbtq_clean, evaluation_minority_clean) |>
  filter(
    evaluation_lgbtq_clean != "not mentioned",
    evaluation_lgbtq_clean != "neutral",
    evaluation_minority_clean != "not mentioned",
    evaluation_minority_clean != "neutral",
  ) |>
  unite(col = "Evaluation", evaluation_lgbtq_clean, evaluation_minority_clean, sep = "/") |>
  ggplot(aes(x = date, y = n, colour = Evaluation)) +
  # geom_line() +
  geom_smooth(method = "loess", formula = "y ~ x", se = FALSE) +
  scale_colour_manual(values = c(
    # highlight
    "positive/positive" = "#77c97d",
    "positive/negative" = "#ffbf47",
    "negative/positive" = "#ffe347",
    "negative/negative" = "#fc4e51"
  )) +
  labs(x = NULL, y = NULL, fill = NULL, colour = NULL) +
  theme_minimal()
ggsave("figures/fig5_alt.png", fig5_evaluation_over_time_alt3, width = 16, height = 11, units = "cm")
fig5_evaluation_over_time_alt3
```

# Wrap-up

```{r}
sessionInfo()
Sys.time()
# note how long the script takes to (re-)run
Sys.time() - start
```

